---
title: "Maps in R (!!!)"
author: "Laura Tremblay-Boyer"
date: "5 April 2018"
output: 
 html_document: 
  toc: true
  toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, # shuts down messages
                      warning = FALSE, 
                      cache = TRUE,
                      eval = FALSE) # saves chunk content so it doesn't rerun everytime
```

```{r}
library(tidyverse)
options(stringsAsFactors=FALSE) # turn-off automatic factors
library(maps)
library(mapdata)
library(maptools)
load('Pacific-region-to-keep.RData', verbose=TRUE)
```

# Data import and formatting
In R data objects have extension .RData, then can be saved with `save()` and loaded back in R with `load()`

```{r}
load('lbest-sample.RData', verbose=TRUE) #lbest
```

## Data formatting
```{r}
yft80 <- lbest %>% filter(yy==1980) %>%
  group_by(lond, latd) %>% 
  summarise(catch=sum(yft_no), 
            effort=sum(hhooks)) %>%
  mutate(cpue=catch/effort)
head(yft80)
```

# Making the map 
The ggplot function to do heatmaps is called `geom_tile()`
```{r}
ggplot() + 
  geom_tile(data=yft80, aes(x=lond, y=latd, fill=cpue), colour='white') +
  theme_bw() + 
  coord_equal() + 
  scale_fill_distiller(palette='Spectral')
```

## Adding continents to my map 
`world2Hires` is a high resolution database of continent outlines
The `2` means it's the Pacific-centred version
```{r}
all_states <- map_data('world2Hires')
reg2keep <- c(reg2keep, 'Vietnam')
states <- all_states %>% filter(region %in% reg2keep)
```

Now that I have formatted by continent dataset, I can add it 
as a layer to ggplot: 
```{r}
ggplot() + 
  geom_tile(data=yft80, aes(x=lond, y=latd, fill=cpue), colour='white') +
  geom_polygon(data=states, aes(x=long, y=lat, group=group), fill='grey95', colour='navy') + 
  theme_bw() + 
  coord_equal(xlim=c(100,280), ylim=c(-50,50)) + 
  scale_fill_distiller(palette='Spectral')
```
