---
title: "All the ggplots"
author: "Laura Tremblay-Boyer"
date: "5 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

```{r}
options(stringsAsFactors = FALSE) # turn-off character as factors
library(tidyverse)
library(lubridate)
## Loading my data
vanuatu.dat <- read.csv('Datasets/Vanuatu report 65.csv')
van1 <- dplyr::select(vanuatu.dat, 
                      trip_date, sp_name, sp_code, 
                      catch_n=sp_n, catch_kg=sp_kg)
focussp <- "DEEPWATER RED SNAPPER"

van1$datefrm <- mdy_hms(van1$trip_date)
van1$Month <- month(van1$datefrm)
van1$Year <- year(van1$datefrm)

vanfilt <- van1 %>% filter(sp_name==focussp)
moncatch <- vanfilt %>% 
  group_by(Month) %>% 
  summarise(catch=mean(catch_kg), 
            catchsd=sd(catch_kg), 
            maxn=max(catch_n))
```

# Making a line plot in ggplot
```{r lineplot}
theme_set(theme_bw())
ggplot(moncatch, aes(x=Month, y=catch), size=2) + 
  geom_line(colour='dodgerblue') + geom_point()

```
Common error message: `invalid argument to unary operator`

# Making a panel line plot in ggplot 
Show the monthly catch by year, for 2015, 2016 and 2017
```{r}
names(vanfilt)
ymoncatch <- vanfilt %>% 
  group_by(Month, Year) %>% 
  summarise(catch=mean(catch_kg), 
            catchsd=sd(catch_kg), 
            maxn=max(catch_n)) %>%
  filter(Year %in% 2015:2017)
unique(ymoncatch$Year)
ymoncatch
```

Now let's make the plot:
To make panels, use `facet_wrap()`
```{r ymonpanel}
theme_set(theme_bw())
ggplot(ymoncatch, aes(x=Month, y=catch, colour=as.factor(Year))) + geom_point() + geom_line() + facet_wrap(~Year)
```

Plotting with panels in rows intead, use the function 
`facet_grid()` (can also use `facet_wrap` with `ncol=1`)
```{r ymonpanel2}
theme_set(theme_bw())
ggplot(ymoncatch, aes(x=Month, y=catch, colour=as.factor(Year))) + geom_point() + geom_line() + 
  facet_grid(Year~.)
```

Plotting with a barplot instead of lines 
```{r}
ggplot(ymoncatch, 
       aes(x=Month, y=catch, fill=as.factor(Year))) + 
  geom_bar(stat='identity', position='dodge')
```

Mixing up barplot with panels
```{r}
ggplot(ymoncatch, 
       aes(x=Month, y=catch, fill=as.factor(Year))) + 
  geom_bar(stat='identity') + 
  facet_wrap(~Year, scale='free_y') + theme_bw() +
  scale_fill_manual(values=c('dodgerblue', 'turquoise', 'indianred'))
```

# Exercise
From the Vanuatu dataset, make a month-panel bar plot of the annual average catch (kg) for dogtooth tuna (all years before the future).

## Filter dataset for dogtooth tuna, and calculate average 
catch by year and month 
```{r}
# find out how dogtooth tuna is referred to in sp_name column
# unique(grep('TUNA',van1$sp_name, value=TRUE))
dogt <- van1 %>%
  filter(sp_name == 'DOGTOOTH TUNA') %>% 
  group_by(Year, Month) %>% 
  summarise(avgcatch=mean(catch_kg)) %>% 
  filter(Year < 2019)
```

Make the plot now that I have created my dataset `dogt`
```{r}
ggplot(dogt, aes(x=Year, y=avgcatch)) + 
  geom_bar(stat='identity') + facet_wrap(~Month) +
  theme_bw()
```

